Exploratory Data Analysis of White Wines
by Ashutosh Singh
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data = read.csv('wineQualityWhites.csv')

#names of variables
names(data)

#dimensions of data
dim(data)

#variables information
str(data)

#convert quality data as factored
data$quality <- as.factor(data$quality)

```
Only the quality column which tells about the average quality of Wines can be used a  categorical column in first pass. Others look like continuous values. Let's see each of them one by one.

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}

#Quality of Wines
qplot(quality,data=data, geom="histogram")

summary(data$quality)
```
Most the wines are of average quality. So we have few data points to delve into qualities. 

```{r echo=FALSE}
qplot(fixed.acidity, data = data, geom="histogram")
summary(data$fixed.acidity)

qplot(volatile.acidity, data = data, geom="histogram")
summary(data$volatile.acidity)

qplot(citric.acid, data = data, geom="histogram")
summary(data$citric.acid)

qplot(chlorides, data = data, geom="histogram")
summary(data$chlorides)

qplot(free.sulfur.dioxide, data = data, geom="histogram")
summary(data$free.sulphur.dioxide)


qplot(density, data = data, geom="histogram")
summary(data$density)


qplot(pH, data = data, geom="histogram")
summary(data$pH)


qplot(sulphates, data = data, geom="histogram")
summary(data$sulphates)


qplot(alcohol, data = data, geom="histogram")
summary(data$alcohol)

```


# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
