##Exploratory Data Analysis of White Wines
##Ashutosh Singh

------


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

#install libraries
#install.packages("memisc")



#load libraries
library(ggplot2)
library(memisc)
library(GGally)
library(corrplot)
library(gridExtra)
source("https://raw.githubusercontent.com/briatte/ggcorr/master/ggcorr.R")




###some useful functions 
## Remove outliers 
## source : http://stackoverflow.com/a/4788102

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}


```

Looking at the names and values of data

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
data = read.csv('wineQualityWhites.csv')

#names of variables
names(data)

#dimensions of data
dim(data)

#na values 
rows <- is.na(data)
sum(rows)

#variables information
str(data)

#convert quality data as factored
data$quality <- as.factor(data$quality)

```
Only the quality column which tells about the average quality of Wines can be 
used a  categorical column in first pass. Others look like continuous values.
Let's see each of them one by one.

# Univariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

#Quality of Wines
qplot(quality,data=data, geom="histogram")

summary(data$quality)
```
Most the wines are of average quality. So we have few data points to delve into
qualities. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 <- qplot(fixed.acidity, data = data, geom="histogram")
summary(data$fixed.acidity)

p2 <- qplot(volatile.acidity, data = data, geom="histogram")
#summary(data$volatile.acidity)

p3 <- qplot(citric.acid, data = data, geom="histogram")
#summary(data$citric.acid)

p4 <- qplot(chlorides, data = data, geom="histogram")
#summary(data$chlorides)

p5 <- qplot(free.sulfur.dioxide, data = data, geom="histogram")
#summary(data$free.sulphur.dioxide)


p6 <- qplot(density, data = data, geom="histogram")
#summary(data$density)


p7 <- qplot(pH, data = data, geom="histogram")
#summary(data$pH)


p8 <- qplot(sulphates, data = data, geom="histogram")
#summary(data$sulphates)


p9 <- qplot(alcohol, data = data, geom="histogram")
#summary(data$alcohol)

p10 <- qplot(residual.sugar, data=data, geom="histogram")


#arrange the plots 
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,ncol=3)


```

The residual sugar histogram is skewed so plotting it in log transformation we
get 
```{r echo=FALSE, message=FALSE, warning=FALSE,Residual_Sugar_transform} 

qplot(data = data, x = residual.sugar , geom="histogram")+
  scale_x_log10()

```

There exists a bimodal distribution. Residual sugar is what is left when the 
fragmentation stops. 


# Univariate Analysis

### What is the structure of your dataset?
The data contains 4898 observations of white wine and their physiochemical 
analysis as features. A total of 11 features are used and the quality variable 
is used as output which also serves as a categorical variable.

All features except quality are continuous.The quality feature domain is from 
0 to 10 but the range is available from 3 to 9. The quality of wine is 
subjective so  average was used to depict quality by atleast 3 observers.

### What is/are the main feature(s) of interest in your dataset?
The quality feature rises to the top . Also since I have never tasted wine and 
as an engineer I thought of predicting the quality of wine by its features. 
Residual sugar also seems interesting as this what makes wines taste sweeter.
Also content of residual sugar is maintained and the sulphides are  added to
preserve the sugar content from further fermentation. 

Last I think most important is alcohol for obvious reasons. 

### What other features in the dataset do you think will help support your
investigation into your feature(s) of interest?

I think pH value and density features should be explored more in association 
with the quality of wines

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions?
Did you perform any operations on the data to tidy, adjust, or change the form
of the data? If so, why did you do this?
I investigated the residual sugar feature since the data looks skewed in the
first pass. I applied the log transformation on it which clearly shows a bimodal
relationship between the wines and the sugar after fermentation. 
The plot shows that mostly the sugar is determined by keeping it low or high.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}


#ggpairs(data = data,columns=2:13)

#plotting the correlation matrix
#cormat <- cor(data[2:12], data[2:12])
ggcorr(data=data[2:12],geom="tile", label=TRUE, hjust=0.9,
       layout.exp=3 , nbreaks = 5, label_round =2) +
  ggplot2::labs(title="Correlation Values between features")

```

For a quick overview of the data I plotted a scatterplot matrix of the
variables to look at the correlation and the distribution with other variables.

Following strong correlations are found 

 - free sulphur dioxide and total sulphur dioxdie (0.616)
 - alcohol and quality (0.436)
 - fixed acidity and pH (0.426)
 - residual sugar and density (0.839)

The scatter plot also shows many outliers here and there so I need to remove
the
outliers before further processing.The correlation between density and pH 
is **-0.9**.


```{r echo=FALSE, message=FALSE, warning=FALSE, Remove_Outliers}

#scttter plot of density with pH
ggplot(data = data, aes(x = density, y = pH))+
  geom_point(alpha=1/4)+
  stat_smooth(method="lm", col="red") +
  ggtitle("density and pH  with outliers")

#density shows some outliers so first remove it
data$density <- remove_outliers(data$density)

ggplot(data = data, aes(x = density, y = pH))+
  geom_point(alpha =1/4)+
  stat_smooth(method="lm", col="red")+
  ggtitle("density and pH with outliers removed")

```

From the above pairs matrix a relationship between density and residual sugar 
looks linear. The correlation between them is **0.83**
Plotting them again 

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = data , aes(x = density, y = remove_outliers(residual.sugar))) +
  geom_point(alpha=1/4) +
  stat_smooth(method="lm", color="darkred")
  
#data$chlorides <- remove_outliers(data$chlorides)

```

Since our categorical value and also output is only quality. Lets take a look 
at how the values of different features relate with quality.

```{r echo=FALSE, message=FALSE, warning=FALSE, Feature_Box_Plots}

ggplot(data,aes(factor(quality),fixed.acidity))+geom_boxplot() +xlab("quality") +ylab("fixed acidity")
```

The y axis spread is too much due to the outliers but we get the idea. Fixed
acidity does is not discrminative enough to change according to quality. Similar
will be the case for volatile acidity as fixed acidity and volatile acidity are
linearly correlated. We also see the same trend for Citric Acid

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Acidity_Comparison}
ggplot(data,aes(factor(quality),volatile.acidity)) + geom_boxplot() +
  xlab("quality") + ylab("volatile acidity") +
  ggtitle("Volatile Acidity and Quality")


ggplot(data,aes(factor(quality),citric.acid)) + geom_boxplot() +
  xlab("quality") + ylab("Citric Acid") +
  ggtitle("Citric acid and Quality")


```

Looking at the relationship between residual sugar and qaulity

```{r echo=FALSE, message=FALSE, warning=FALSE, Residual_Quality}

ggplot(data,aes(x=quality, y= remove_outliers(residual.sugar)))+
  geom_boxplot()+
  ylab("Residual Sugar")+
  ggtitle("Residual sugar and quality")

```

It looks like low residual sugar and wine quality are correlated but we also 
get almose same values of residual sugar for wines rated poor. It is inconclusive
to use residual sugar for discriminating wines.The median value is low than others

Moving on to quality and chlorides

```{r echo=FALSE, message=FALSE, warning=FALSE, Chlorides_and_Quality}

pcl1 <- ggplot(data,aes(x=quality, y= chlorides))+
  geom_boxplot()+
  ylab("Chloride")+
  ggtitle("Chlorides and quality")

pcl2 <- ggplot(data,aes(x=quality, y= remove_outliers(chlorides)))+
  geom_boxplot()+
  ylab("Chlorides (no outliers)")+
  ggtitle("Chlorides and quality")+
  stat_summary(fun.y=mean, color="red",geom="point",shape=18,size=3)
  

grid.arrange(pcl1,pcl2,ncol=2)

```

I have removed the outliers from the chlorides data as they were skewing the
plot and we may be missing the big picture.The red point shows the mean values.
From the box plot it is clear that high quality wines have low chlorides.
We will explore this aspect in detail later on.

The next feature is Free sulfur dioxide. Since *free sulfur dioxide* and
*total sulfur dioxide* are correlated with almost linear relationship,looking
at them show similar plots

```{r echo=FALSE, message=FALSE, warning=FALSE, Sulfur_and_quality}

ps1 <- ggplot(data,aes(x=quality, y=free.sulfur.dioxide)) +
  geom_boxplot()

ps2 <- ggplot(data,aes(x=quality, y=total.sulfur.dioxide)) +
  geom_boxplot()

grid.arrange(ps1,ps2,ncol=2)

```

The plot looks almost same except for scale which is obvious and there is also 
strong correlation between them as shown in plot below.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data,aes(x=total.sulfur.dioxide, y = free.sulfur.dioxide))+
  geom_point(alpha=1/10, position="jitter")+
  stat_smooth(method = "lm")

```


Now looking at quality with pH, there is a meek indication of high pH as 
compared to othes for high quality wines.

```{r echo=FALSE, message=FALSE, warning=FALSE}

  ggplot(data,aes(x=quality, y=pH))+
    geom_boxplot()+
    ggtitle("pH and Quality")

```

But no such patter emerges when we compare sulphates with the quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}

  ggplot(data,aes(x=quality, y=sulphates))+
    geom_boxplot() +
    ggtitle("Sulphates and Quality ")
  

```

Finally we have alcohol with quality. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

  ggplot(data,aes(x=quality, y=alcohol))+
    geom_boxplot() +
    ggtitle("Alcohol and Quality ")

```



These box plots above give us an average idea of what features change according
to quality. Some specific mentions are

* chlorides have low quantity with high quality
* density is also low for high quality wines
* alcohol is high for high quality wines

These comparisons give us a bird-eye view of how individual features impact the
quality of the wines. Now lets take an overview of how these features
affect each other.



```{r echo=FALSE, message=FALSE, warning=FALSE}

data$clean.chlorides <- remove_outliers(data$chlorides)

ggplot(data= data,aes(clean.chlorides,color=quality,fill=quality,group=quality)) + 
       geom_density() +
       ggtitle("Density histogram of chlorides")+
       xlab("Chlorides")
```

We get a clear picture that high quality rated wines have less chlorides.But 
the correlation is not straight. Low quality rated wines (quality 3) also have
less chlorides than the average. There may be other factors. 

Taking another look at density feature and quality

```{r echo=FALSE, message=FALSE, warning=FALSE}

data$clean.density <- remove_outliers(data$density)

qplot(data = data, x= factor(quality), y = clean.density )+
  geom_boxplot()+
  ylab("Density")+
  xlab("Quality")+
  ggtitle("Density and Quality")

table(data$quality)

ggplot(data= data,aes(clean.density,color=quality,fill=quality,group=quality)) +
  geom_density()+
  ggtitle("density histogram of wine \"Density\"")+
  xlab("Wine Density")

```

From the plot and the values we have less data for high quality rate wines :
only 1060 but it is evident that high-rated wines have less density than the
low rated wines. 


And now taking a look at the boxplot of alcohol and quality

```{r echo=FALSE, message=FALSE, warning=FALSE}

data$clean.alcohol <- remove_outliers(data$alcohol)

qplot(data=data, x = factor(quality), y = clean.alcohol)+
  geom_boxplot()

ggplot(data= data,aes(clean.alcohol,color=quality,fill=quality,group=quality)) +
  geom_density()

```

The intuition was right the alcohol content of the high-rated wines is greater
than in low-rated wines.

There seems to be some relation between pH and fixed acidity from the plot
matrix above

```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(data=data,x = pH, y = fixed.acidity) + 
  ggtitle("pH and fixed.acidity")
```

Plotting with noise to remove the overplotting.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = data,aes(x = pH, y = fixed.acidity))+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  ggtitle("pH and fixed.acidity with linear model") +
  stat_smooth(se=FALSE, method=lm,col="red")

```

It looks like a negative correlation between pH and **fixed.acidity** which is 
trivial as pH is an indicator of the acidity.The above plot also shows 
a linear regression line. The correlation between pH and fixed acidity is 
**-0.43**

```{r echo=FALSE, message=FALSE, warning=FALSE}

data$clean.totol.sulfur.dioxide <- remove_outliers(data$total.sulfur.dioxide)

ggplot(data=data,aes(x=clean.totol.sulfur.dioxide,
                     remove_outliers(residual.sugar)))+
                     geom_point(position="jitter", alpha=0.5, size=0.75) +
                    xlab("total sulfur dioxide (no outliers)")+
                    ylab("residual sugar (no outliers)")
                    stat_smooth(method="lm", color="darkred")
  

```

There are too many points on the lower side of the residual sugar even after
removing the outliers but on the upper side it shows a linear relationship.
The correlation between residual sugar and total sufur dioxide is **0.41**

Same relationships can also be found in between 

* alcohol and residual.sugar (correlation **-0.46**)
* alcohol and total.sulfur.dioxide (correlation **-0.45**)

Plots are below

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data=data, aes(x = alcohol, y = residual.sugar))+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  stat_smooth(method="lm", col="red")

ggplot(data=data, aes(x = alcohol, y = clean.totol.sulfur.dioxide))+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  stat_smooth(method="lm", col="red")

```




# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the 
investigation. How did the feature(s) of interest vary with other features 
in the dataset?
Based on the correlation coefficients between the various pairs the most strong
relationship is found beween the following pairs

Sr No.    Feature1                  Feature2             Correlation
------    --------                  ---------            -----------
1.        pH                        fixed.acidity         -0.4258
2.        total.sulfur.dioxide      residual.sugar        0.4014
3.        alcohol                   sugar                 -0.4506
4.        achohol                   total.sulfur.dioxide  -0.4488


I have not taken the quality in calculating the correlation as I just wanted
to know the correlation between the features themselves.

### Did you observe any interesting relationships between the other features
(not the main feature(s) of interest)?

Yes, as shown in the plots and the table above, I found some relationships
which were not evident directly. Plotting variables helped to look into these 
in more detail. 

The relationship between the pH value and the fixed.acidity is trivial as pH 
value is a representation of the acidity. But the relationship between the  
total sulfurdioxide and the residual sugar , alcohol and residual sugar may 
help us in explaining the variance which will be explored in the next section.  


### What was the strongest relationship you found?
The strongest relationship found was between the free sulfur dioxide and the
total sulfur dioxide. This is trivial as both are interdependent. Also their 
correlation doesn't matter because we one is a part of other and hence they
will also lead to multi collinearity problems if we use both of them in a
linear model. 

Other than that the strongest relationship is a negative correlation between
the density and residual sugar.

# Multivariate Plots Section

Before starting the multivariate analysis I want to convert more features into 
categorical values. These are

* **Alcohol** It is cut into 7 factors from 8 to 15 with an interval of 1

```{r echo=FALSE, message=FALSE, warning=FALSE}

data$alcohol.category<-cut(data$alcohol, seq(8,15,1), right=FALSE,labels=c(1:7))
data$quality.cat <- cut(as.numeric.factor(data$quality),breaks = c(0,4,6,10), labels=c("poor","average","good"))

summary(data$alcohol.category)

```

* residual.sugar

Since the residual.sugar is bimodal on the log10 scale so I convert it into 
the orders of low and high based on its median.

```{r echo=FALSE, message=FALSE, warning=FALSE}

summary(data$residual.sugar)
#median is 5.200 so cut it into high and low with the median
data$residual.sugar.cat <- cut(data$residual.sugar, c(0,median(data$residual.sugar),max(data$residual.sugar)),labels=c("low","high"))
summary(data$residual.sugar.cat)

```

* quality

quality is cut into 3 categories of **poor, average and good**

```{r ehco=FALSE, warning=FALSE, message=FALSE}

summary(data$quality.cat)

```

Since density and alcohol are co-related with the quality. lets take a look 
at them by quality

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = data, aes(x = density, y = alcohol, color=quality.cat)) + 
  geom_point(alpha=1/3, position="jitter") +
  stat_smooth(method = "lm")+
  ggtitle("Alcohol v/s Density colored by quality")+
  scale_colour_manual(values=c("red", "#6666dd", "#00ff00"))


```

There are some points at the top left telling a better quality wine is with
more alcohol and less density. Also the linear model lines tell that we get 
better quality at low density and high alcohol level. The good quality wines 
are separable at these levels.


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

qplot(data =data, x= alcohol, y = density, color=residual.sugar.cat) +
  stat_smooth(method = "lm")
```

This plot shows the relationship between the density, alcohol and the residual 
sugar. Following relatiionships are evident 

* for more alcohol we have less sugar content and also the density tends to be low
* high density correlates to high sugar content. 
* There is negative relationship between the the alcohol content and density 
of the wines

```{r echo=FALSE, message=FALSE, warning=FALSE}
  
ggplot(data = data, aes(y = density,
                        x= remove_outliers(residual.sugar),                                                   color=quality.cat) ) +
      geom_point(position="jitter", alpha=1/2)+
      stat_smooth(method="lm")+
      xlab("Residual Sugar")

```

This plot shows density and residual sugar with points colored buy the quality
of wine. The plot is mostly green as there are more number of average wines but
the purple color at the botom explains some of the variance in the quality with
lower density for same abount of residual sugar for good quality wines. 

### Now making a predictive model

In the first iteration I tool all the features in the model 

```{r}



data$quality.num <- as.numeric.factor(data$quality)
m1 <- lm(quality.num ~ fixed.acidity , data=data)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + citric.acid)
m4 <- update(m3, ~ . + residual.sugar)
m5 <- update(m4, ~ . + chlorides)
m6 <- update(m5, ~ . + free.sulfur.dioxide)
m7 <- update(m6, ~ . + total.sulfur.dioxide)
m8 <- update(m7, ~ . + density)
m9 <- update(m8, ~ . + pH)
m10 <- update(m9, ~ . + sulphates)
m11 <- update(m10, ~ . + alcohol)

mtable(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11)

```

The R-square value goes to a maximum of 0.286, not so good. Also neither of 
the features intercept and p-value looks significant. More  can be explored 
in the model are as the features are added in sequence basis and not in a 
calculated  manner. 
The R-squared value may be increased by few percentage points but that will
not take more work that the accuracy gained.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the
investigation. Were there features that strengthened each other in terms of
looking at your feature(s) of interest?

I used the multivariate analysis to quantify more data from the bi-variate
analysis. The plots are redone with taking some feature as a colored factor.
Relationships like as density increases the residual.sugar also increases.
There is also a very strong correlation between the density and residual.sugar .


### Were there any interesting or surprising interactions between features?

The most interesting interaction comes between the density, alcohol and 
residual sugar. With the decrease density the alcohol content increases but
the residual sugar in the wines decrease. There is clear line of separation 
between them based on residual sugar. 

The most important part that I found by the uni-variate, bi-variate and 
multi-variate analysis is that there is too much overlap between the features
and nothing can be conclusively said about the wine quality from its chemical
composition. 


### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

Yes, 11 linear models were created with adding features one by one. 
The highest r-squared value came to be  **0.286** when all the features are 
incorporated. The variance explaination is not much and hence it cannot be 
said that we have a good model for prediction of wine quality. 

#### Strengths:
* The variance explaination of the model increased with every added feature
and hence it can be said that each feature contributes in someway towards 
predicting the wine quality.

* The model also explains which feature are more important than others 
like : alcohol, residual.sugar and density. 


#### Limitations

* The model doesn't explain much of the variance for the quality of wine. 
The r-squared value is just 0.286.

* The reason for the above may be because we don't have enough data.
Most of the data is of average quality and many features  overlap between
the good quality and average quality hence a consensus cannot be found
between the important features and their values. 

* There is strong correlation between features such as free.sulfur.dioxide and
the total.sulrfur dioxide ; density and residual.sugar. This also leads to
problem of multicollinearty. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

p1 <- qplot(data = data, x = residual.sugar, xlim = c(0,50)) +
  ylab("Number of observations") +
  xlab("Residual Sugar") +
  ggtitle("Histogram of Residual Sugar content in wine")+
  theme(text = element_text(size=14))
  

p2 <- qplot(data=data, x = residual.sugar, binwidth = log10(1.05) , xlim=c(0,50))+
  geom_vline(xintercept = mean(data$residual.sugar), color="red", show_guide=TRUE)+
  ylab("Number of Observations")+
  scale_x_log10("Residual Sugar (log10)")+
  ggtitle("Histogram of Residual sugar content in wine") +
  scale_colour_manual(values=c('red'='median', 'black'='none'))+
  theme(text = element_text(size=14))


grid.arrange(p1,p2,ncol=2)
```

### Description One
This plot is the most interesting plot that I found. Both the plots show the
histogram of residual sugar content in the wines data. But the plot on the 
left is on normal scale and that on right is on logarithmic scale. The first
plot shows that the residual sugar data is left skewed and most of the values
are less than 20 in decreasing order at every level. 
The second plot shows the same data but on the logarithmic scale and this is
where the hidden feature comes into the picture.  The second plot is bimodal.
The *median value of the feature is shown in red line*. The bimodal plot starts 
sometime before the median and it roughly divides the two distributions in equal
part. 

This tells us that there is clear distinction in the wines data based on the
residual sugar and it is explored more in the next plot. 



### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(data =data, aes(x= alcohol, y = density, color=residual.sugar.cat)) +
  geom_point(alpha=1/2) +
  ggtitle("Distibution of density and alcohol classified by residual sugar.")+
  xlab("Alcohol") +
  ylab("Density") + 
  scale_colour_discrete(name="Residul Sugar",
                        labels=c("Low (<0.52)","High (>0.52)")) +
  stat_smooth(method="lm")

```

### Description Two

This plot shows the three-way relationship between density, alcohol and 
residual sugar. The idea of breaking the residual sugar into a factor comes 
from the first plot as it shows the bimodal relation ship. The median is
chosen as the breakpoint.

This plot shows the strong relationship between the density-alcohol and
density-residual sugar.  The density decreases with the increase in alcohol
level and high density leads to high level of sugar. The plot also build upon
the hypothesis of previous plot that in the wines data there is clear 
distinction by the residual sugar level.


### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
rows <- is.na(data$density)
data <- data[!rows,]

#calculate correlation matrix. 

cormat <- cor(data[2:12], data[2:12])


#corrplot(cormat,method="number",type="lower", diag=FALSE ,title="Correlation of Features",add=FALSE, mar=c(0,0,1,0),col="black")
ggcorr(data=data[2:12],geom="tile", label=TRUE, hjust=0.9,
       layout.exp=3 , nbreaks = 5, label_round =2) +
  ggplot2::labs(title="Correlation Values between features")

```

### Description Three
This plot/matrix is the correlation matrix between the different features of 
the wines data set. The closer to zero correlations are in white hence less 
visible than the correlations which are closer to extremes ( negatives in red
and positives in blue) . Only taking a look at the matrix tells us about highly
correlated features like 

* density and sugar (0.83)
* density and alcohol (-0.81)
* alcohol and residual.sugar (0.46)

This plot/matrix brings out the best of correlations and displays it in easier
to read formats.

------

# Reflection

This was an intense project compared previous. The data was just thrown and I
needed to find the patterns. I chose the White Wines dataset as I didn't had 
enough time to dig into other datasets and this was also a tidy one so I thought
it will be easier. I started with exploring through the features and and then 
finding relationship between them to making a model to predict the wine quality.
Some features are correlated strongly and many are not. The prediction model
doesn't do well as the features are variable throughout the quality level.
There is also a strong bias towards the average quality wines as the data for 
the high and poor quality is considerably less.

The future work in this can be done with some more data about the good quality 
wines. Using a better and calculated approach to feature selection can improve
the model with less features also. Also I haven't used the dplyr, etc. packages 
which can be used to make more intricate analysis.

In the end I started with zero knowledge of wines but now I know what components
affects wine 
quality. The knowledge is not sufficient to be wine connoisseur but it's better
than a blind guess. 



# References 
1. https://en.wikipedia.org/wiki/White_wine
2. http://www.r-bloggers.com/from-continuous-to-categorical/
3. Thinking fast and slow , Chapter 21 : Intutions vs Formulas ; Daniel Kahneman
4. https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html
5. https://en.wikipedia.org/wiki/White_wine
6. http://docs.ggplot2.org/0.9.3.1/guide_legend.html
7. http://stackoverflow.com/a/4788102
8. https://briatte.github.io/ggcorr/