Exploratory Data Analysis of White Wines
by Ashutosh Singh

========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
#install.packages("memisc")
library(memisc)


###some useful functions 
## Remove outliers 
## source : http://stackoverflow.com/a/4788102

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}


```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data = read.csv('wineQualityWhites.csv')

#names of variables
names(data)

#dimensions of data
dim(data)

#variables information
str(data)

#convert quality data as factored
data$quality <- as.factor(data$quality)

```
Only the quality column which tells about the average quality of Wines can be used a  categorical column in first pass. Others look like continuous values. Let's see each of them one by one.

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}

#Quality of Wines
qplot(quality,data=data, geom="histogram")

summary(data$quality)
```
Most the wines are of average quality. So we have few data points to delve into qualities. 

```{r echo=FALSE}
library(gridExtra)

p1 <- qplot(fixed.acidity, data = data, geom="histogram")
summary(data$fixed.acidity)

p2 <- qplot(volatile.acidity, data = data, geom="histogram")
#summary(data$volatile.acidity)

p3 <- qplot(citric.acid, data = data, geom="histogram")
#summary(data$citric.acid)

p4 <- qplot(chlorides, data = data, geom="histogram")
#summary(data$chlorides)

p5 <- qplot(free.sulfur.dioxide, data = data, geom="histogram")
#summary(data$free.sulphur.dioxide)


p6 <- qplot(density, data = data, geom="histogram")
#summary(data$density)


p7 <- qplot(pH, data = data, geom="histogram")
#summary(data$pH)


p8 <- qplot(sulphates, data = data, geom="histogram")
#summary(data$sulphates)


p9 <- qplot(alcohol, data = data, geom="histogram")
#summary(data$alcohol)

p10 <- qplot(residual.sugar, data=data, geom="histogram")


#arrange the plots 
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,ncol=3)


```

The residual sugar histogram is skewed so plotting it in log transformation we get 
```{r Residual Sugar transform} 

qplot(data = data, x = residual.sugar , geom="histogram")+
  scale_x_log10()

```
There exists a bimodal distribution. Residual sugar is what is left when the fragmentation stops. 




# Univariate Analysis

### What is the structure of your dataset?
The data contains 4898 observations of white wine and their physiochemical analysis as features. A total of 11 features are used and the quality variable is used as output which also serves as a categorical variable.

All features except quality are continuous.The quality feature domain is from 0 to 10 but the range is available from 3 to 9. The quality of wine is subjective so an average was used to depict quality by atleast 3 observers.

### What is/are the main feature(s) of interest in your dataset?
The quality feature rises to the top as I have only seen in movies that they get the quality of wine by some ritual. Also since I have never tasted wine and as an engineer I thought of predicting the quality of wine by its features. 
Residual sugar also seems interesting as this what makes wines taste sweeter. Also content of residual sugar is maintained and the sulphides are aslo added to preserve the sugar content from further fermentation. 

Last I think most important is alcohol for obvious reasons. 

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I think pH value and acidity features should be explored more in association with the quality go wines

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
I investigated the residual sugar feature since the data looks skewed in the first pass. I applied the log transformation on it which clearly shows a bimodal relationship between the wines and the sugar after fermentation. 
The plot shows that mostly the sugar is determined by keeping it low or high.

# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
library(GGally)
library(corrplot)

#ggpairs(data = data,columns=2:13)

#plotting the correlation matrix
cormat <- cor(data[2:12], data[2:12])
corrplot(cormat, method="number")

```

For a quick overview of the data I plotted a scatterplot matrix of the variables to look at the correlation and the distribution with other variables. 

Follwing strong correlations are found 

 - free sulphur dioxide and total sulphur dioxdie (0.616)
 - alcohol and quality (0.436)
 - fixed acidity and pH (0.426)
 - residual sugar and density (0.839)

The scatter plot also shows many outliers here and there so I need to remove the outliers before further processing.


```{r echo=FALSE, Remove_Outliers}

#scttter plot of density with pH
qplot(data = data, x = density, y = pH)+
  geom_point()

#density shows some outliers so first remove it
data$density <- remove_outliers(data$density)

qplot(data = data, x = density, y = pH)+
  geom_point()

```

From the above plot matrix a relationship between density and residual sugar looks linear. Plotting them again 

```{r echo=FALSE}

qplot(data = data , x = density, y = residual.sugar)+
  geom_point() 
  
#data$chlorides <- remove_outliers(data$chlorides)

qplot(data=data,y=density,x= sulphates)+  
  geom_boxplot()

```

Since our categorical value and also output is only quality. Lets take a look at how the values of different features spread with quality

```{r echo=FALSE, Feature_Box_Plots}

pf1 <- ggplot(data,aes(factor(quality),fixed.acidity))+geom_boxplot() +xlab("quality")
pf2 <- ggplot(data,aes(factor(quality),volatile.acidity))+geom_boxplot()+xlab("quality")
pf3 <- ggplot(data,aes(factor(quality),citric.acid))+geom_boxplot()+xlab("quality")
pf4 <- ggplot(data,aes(factor(quality),residual.sugar))+geom_boxplot()+xlab("quality")
pf5 <- ggplot(data,aes(factor(quality),chlorides))+geom_boxplot()+xlab("quality")
pf6 <- ggplot(data,aes(factor(quality),free.sulfur.dioxide))+geom_boxplot()+xlab("quality")
pf7 <- ggplot(data,aes(factor(quality),total.sulfur.dioxide))+geom_boxplot()+xlab("quality")
pf8 <- ggplot(data,aes(factor(quality),density))+geom_boxplot()+xlab("quality")
pf9 <- ggplot(data,aes(factor(quality),pH))+geom_boxplot()+xlab("quality")
pf10 <- ggplot(data,aes(factor(quality),sulphates))+geom_boxplot()+xlab("quality")
pf11 <- ggplot(data,aes(factor(quality),alcohol))+geom_boxplot()+xlab("quality")

grid.arrange(pf1,pf2,pf3,pf4,pf5,pf6,pf7,pf8,pf9,pf10,pf11,ncol=3)

```

These box plots give us an average idea of what features change according to quality. Some specific mentions are

* chlorides have low quantity with high quality
* density is also low for high quality wines
* alcohol is high for high quality wines

I am still interested in residual sugar due to its bimodal histogram on the log axis.So instead of quality lets look at the scatter plots of residual sugar with other features.

Looking into them in bigger values

```{r echo=FALSE}

data$clean.chlorides <- remove_outliers(data$chlorides)

p1<-qplot(data = data, x = factor(quality), y = clean.chlorides)+
  geom_boxplot() + xlab("quality") +
  stat_summary(fun.y=mean, color="red", hape=18,geom="point", size=3,show_guide = TRUE)

p2 <- qplot(data=data, x = factor(quality), y = chlorides)+ geom_boxplot()

grid.arrange(p2,p1, ncol=2)

ggplot(data= data,aes(clean.chlorides,color=quality,fill=quality,group=quality)) + geom_density()

```

We get a clear picture that high quality rated wines have less chlorides.But the correlation is not straight. Low quality rated wines (quality 3) also have less chlorides than the average. There may be other factors. 

Taking another look at density and quality

```{r echo=FALSE}

data$clean.density <- remove_outliers(data$density)

qplot(data = data, x= factor(quality), y = clean.density )+
  geom_boxplot()

table(data$quality > 6)

ggplot(data= data,aes(clean.density,color=quality,fill=quality,group=quality)) +
  geom_density()


```
From the plot and the values we have less data for high quality rate wines : only 1060 but it is evident that high-rated wines have less density than the low rated wines. 


And now taking a look at the boxplot of alcohol and quality

```{r echo=FALSE}

data$clean.alcohol <- remove_outliers(data$alcohol)

qplot(data=data, x = factor(quality), y = clean.alcohol)+
  geom_boxplot()

ggplot(data= data,aes(clean.alcohol,color=quality,fill=quality,group=quality)) +
  geom_density()

```
The intuition was right the alcohol content of the high-rated wines is greater than in low-rated wines.

There seems to be some relation between pH and fixed acidity from the plot matrix above

```{r echo=FALSE}

qplot(data=data,x = pH, y = fixed.acidity) + 
  ggtitle("pH and fixed.acidity")
```
Plotting with noise to remove the overplotting.

```{r echo=FALSE}
ggplot(data = data,aes(x = pH, y = fixed.acidity))+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  ggtitle("pH and fixed.acidity with linear model") + stat_smooth(se=FALSE, method=lm,col="red")

```
It looks like a negative correlation between pH and **fixed.acidity** . The above plot also shows a linear regression line. 

```{r echo=FALSE}

data$clean.totol.sulfur.dioxide <- remove_outliers(data$total.sulfur.dioxide)

ggplot(data=data,aes(x=clean.totol.sulfur.dioxide, residual.sugar))+
  geom_point(position="jitter", alpha=0.5, size=0.75)

```

There are too many points on the lower side of the residual sugar even after removing the outliers but on the upper side it shows a linear relationship.

Same relationships can also be found in between 
* alcohol and residual.sugar
* alcohol and total.sulfur.dioxide

Plots are below

```{r echo=FALSE}

qplot(data=data, x = alcohol, y = residual.sugar)+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  stat_smooth(method="lm", col="red")

qplot(data=data, x = alcohol, y = clean.totol.sulfur.dioxide)+
  geom_point(position="jitter", alpha=0.5, size=0.75)+
  stat_smooth(method="lm", col="red")

```




# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
Based on the correlation coefficients between the various pairs the most strong relation ship is found beween the following pairs

Sr No.    Feature1    Feature2      Correlation
------    --------    ---------     ------------
1.        pH          fixed.acidity  -0.4258
2.        total.sulfur.dioxidex residual.sugar 0.4014
3.        alcohol      sugar          -0.4506
4.        achohol      total.sulfur.dioxide  -0.4488


I have not taken the quality in calculating the correlation as I just wanted to know the correlation between the features themselves.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

Yes, as shown in the plots and the table above, I found some relationships which were not evident directly. Plotting variables helped to look into these in more detail. 

The relationship between the pH value and the fixed.acidity is trivial as pH value is a representation of the acidity. But the relationship between the  total sulfurdioxide and the residual sugar , alcohol and residual sugar may help us in explaining the variance which will be explored in the next section.  


### What was the strongest relationship you found?
The strongest relationship found was between the free sulfur dioxide and the total sulfur dioxide. This is trivial as both are interdependent. Also their correlation doesn't matter because we one is a part of other and hence they will also lead to multi collinearity problems if we use both of them in a linear model. 

Other than that the strongest relationship is a negative correlation between the alcohol and residual sugar.

# Multivariate Plots Section

Before starting the multivariate analysis I want to convert more features into categorical values. These are

* alcohol its cut into 8 to 15 with an iterval of 1
```{r echo=FALSE}

data$alcohol.category<-cut(data$alcohol, seq(8,15,1), right=FALSE,labels=c(1:7))
data$quality.cat <- cut(as.numeric.factor(data$quality),breaks = c(0,4,6,10), labels=c("poor","average","good"))

#There was overplotting in the default case so added the jitter.
qplot(data=data,x=total.sulfur.dioxide, y=alcohol, color=quality.cat)+
geom_point(position="jitter", alpha = 0.5, size=0.75)

```

* residual.sugar

Since the residual.sugar is bimodal on the log10 scale so I convert it into the orders of low and high based on its median.

```{r echo=FALSE}

summary(data$residual.sugar)
#median is 5.200 so cut it into high and low with the median
data$residual.sugar.cat <- cut(data$residual.sugar, c(0,median(data$residual.sugar),max(data$residual.sugar)),labels=c("low","high"))
summary(data$residual.sugar.cat)

```

* quality

quality is cut into 3 catgegories of **poor, average and good**

```{r ehco=FALSE}

data$quality.cat <- cut(as.numeric.factor(data$quality),breaks = c(0,4,6,10), labels=c("poor","average","good"))
summary(data$quality.cat)

```

Since density and alcohol are co-related with the quality. lets take a look at them by quality

```{r echo=FALSE}

qplot(data = data, x = density, y = alcohol, color=quality.cat) + 
  ggtitle("density and alcohol categorized by quality")

```

There are some points at the top left telling a better quality wine is with more alcohol and less density. 


```{r echo=FALSE, Multivariate_Plots}

qplot(data =data, x= alcohol, y = density, color=residual.sugar.cat) +
  stat_smooth(method = "lm")
```
This plot shows the relationship between the density, alcohol and the residual sugar. Following relatiionships are evident 

* for more alcohol we have less sugar content and also the density tends to be low
* high density correlates to high sugar content. 
* There is negative relationship between the the alcohol content and density of the wines

```{r echo=FALSE}
  
qplot(data = data, y = density, x  = residual.sugar, color=quality.cat) +
  geom_point(position="jitter")

```
This plot shows density and residual sugar with points colored buy the quality of wine. The plot is mostly green as there are more number of average wines but the purple color at the botom explains some of the variance in the quality with  lower density for same abount of residual sugar for good quality wines. 

### Now making a predictive model

In the first iteration I tool all the features in the model 

```{r}



data$quality.num <- as.numeric.factor(data$quality)
m1 <- lm(quality.num ~ fixed.acidity , data=data)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + citric.acid)
m4 <- update(m3, ~ . + residual.sugar)
m5 <- update(m4, ~ . + chlorides)
m6 <- update(m5, ~ . + free.sulfur.dioxide)
m7 <- update(m6, ~ . + total.sulfur.dioxide)
m8 <- update(m7, ~ . + density)
m9 <- update(m8, ~ . + pH)
m10 <- update(m9, ~ . + sulphates)
m11 <- update(m10, ~ . + alcohol)

mtable(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11)

```

The R-square value goes to a maximum of 0.286, not so good. Also neither of the features intercept and p-value looks significant. More  can be explored in the model are as the features are added in sequence basis and not in a scientific manner. 
The R-squared value may be increased by few percentage points but that will not take more than than the profit.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

I used the multivariate analysis to quantify more data from the bi-variate analysis. Many plots are redone with taking quality as a colored factor. Relationships like as density increases the residual.sugar also increases. There is also a very strong correlation between the density and residual.sugar .


### Were there any interesting or surprising interactions between features?

The most interesting interaction comes between the density, alcohol and residual sugar. With the decrease density the alcohol content increases but the residual sugar in the wines decrease. There is clear line of separation between them based on residual sugar. 

The most important part that I found by the uni-variate, bi-variate and multi-variate analysis is that there is too much overlap between the features and nothing can be conclusively said about the wine quality from its chemical composition. 


### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

Yes, 11 linear models were created with adding features one by one. The highest r-squared value came to be  **0.286** when all the features are incorporated. The variance explaination is not much and hence it cannot be said that we have a good model for prediction of wine quality. 

#### Strengths:
* The variance explaination of the model increased with every added feature and hence it can be said that each feature contributes in someway towards predicting the wine quality.

* The model also explains which feature are more important than others like : alcohol, residual.sugar and density. 


#### Limitations

* The model doesn't explain much of the variance for the quality of wine. The r-squared value is just 0.286.

* The reason for the above may be because we don't have enough data. Most of the data is of average quality and many features  overlap between the good quality and average quality hence a consensus cannot be found between the important features and their values. 

* There is strong correlation between features such as free.sulfur.dioxide and the total.sulrfur dioxide ; density and residual.sugar. This also leads to multicollinearty. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

p1 <- qplot(data = data, x = residual.sugar, xlim = c(0,50)) + ylab("Number of observations") +
  xlab("Residual Sugar") +
  ggtitle("Histogram of Residual Sugar content in wine")

p2 <- qplot(data=data, x = residual.sugar, binwidth = log10(1.05) , xlim=c(0,50))+
  geom_vline(xintercept = mean(data$residual.sugar), color="red", show_guide=TRUE)+

  ylab("Number of Observations")+
  scale_x_log10("Residual Sugar (log10)")+
  ggtitle("Histogram of Residual sugar content in wine") +
  scale_colour_manual(values=c('red'='median', 'black'='none'))

grid.arrange(p1,p2,ncol=2)
```

### Description One
This plot is the most interesting plot that I found. Both the plots show the histogram of residual sugar content in the wines data. But the plot on the left is on normal scale and that on right is on logarithmic scale. The first plot shows that the residual sugar data is left skewed and most of the values are less than 20 in decreasing order at every level. 
The second plot shows the same data but on the logarithmic scale and this is where the hidden feature comes into the picture.  The second plot is bimodal. The *median value of the feature is shown in read line*. The bimodal plot starts sometime before the median and it roughly divides the two distributions in equal part. 

This tells us that there is clear distinction in the wines data based on the residual sugar and it is explored more in the next plot. 



### Plot Two
```{r echo=FALSE, Plot_Two}
qplot(data =data, x= alcohol, y = density, color=residual.sugar.cat) +
  ggtitle("Distibution of density and alcohol classified by residual sugar.")+
  xlab("Alcohol") +
  ylab("Density") + 
  scale_colour_discrete(name="Residul Sugar", labels=c("Low (<0.52)","High (>0.52)"))

```

### Description Two

This plot shows the three-way relationship between density, alcohol and residual sugar. The idea of breaking the residual sugar into a factor comes from the first plot as it shows the bimodal relation ship. The median is chosen as the breakpoint.

This plot shows the strong relationship between the density-alcohol and density-residual sugar.  The density decreases with the increase in alcohol level and high density leads to high level of sugar. The plot also build upon the hypothesis of previous plot that in the wines data there is clear distinction by the residual sugar level.


### Plot Three
```{r echo=FALSE, Plot_Three}
rows <- is.na(data$density)
sum(rows)
data <- data[!rows,]

cormat <- cor(data[2:12], data[2:12])
corrplot(cormat, method="number")

```

### Description Three
This plot/matrix is the correlation matrix between the different features of the wines data set. The closer to zero correlations are in white hence less visible than the correlations which are closer to extremes ( negatives in red and positives in blue) . Only taking a look at the matrix tells us about highly correlated features like 

* density and sugar (0.83)
* density and alcohol (-0.81)
* alcohol and residual.sugar (0.46)

This plot/matrix brings out the best of correlations and displays it in easier to read formats.

------

# Reflection

# References 
1. https://en.wikipedia.org/wiki/White_wine
2. http://www.r-bloggers.com/from-continuous-to-categorical/