---
title: "Exploratory Analysis of Red Wines Quality"
author: "Ashutosh Singh"
date: "Friday, September 18, 2015"
output:
  html_document:
    highlight: tango
    theme: cerulean
---
###1. Introduction
This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

We are going to work on this to gain experience about the Red Wines composition and how it affects its quality.
First let us load up the data and see some information about it. 

```{r}
#read the csv
wines <- read.csv('wineQualityReds.csv')

#take a look at data 
head(wines)

#a summary of type of data present
str(wines)

```

###2. Univariate Analysis


It looks like there is no FACTORED column here but quality can be used a factor as it is on scale on 1 to 10. 
Converting it to factor
```{r}
wines$quality <- as.factor(wines$quality)

#confirm that it is now represented as factor
str(wines$quality)

```

Plotting a histogram of quality 

```{r}

#load the library
library(ggplot2)
ggplot(data = wines, aes(x = quality))+
  geom_histogram()

```

The plot resembles a  normal distribution. Most of the wines(almost 700) are rated average.

Now let's take a look at alcohol level
```{r}
ggplot(data = wines , aes(x= alcohol))+
  geom_histogram(binwidth = 0.5)

```

The data is a bit skewed on the left side representing that --most-- of the wines have alcohol level of 10% or less.